Approaching Optimal Duplicate Detection in a Sliding Window

### 2020-04-15 appendix.tex
fixes

@@

	When an element $e$ is inserted in the DDF, $e$ is stored and $D[e]$ is incremented (if the key $e$ did not exist in $D$, it is created first, and
	$D[e]$ is set to $1$). In order to keep the number of stored elements to $w$, 
	we discard the oldest element $e_\text{last}$ in $Q$. As we do so, we also decrement $D[e_\text{last}]$, and if $D[e_{last}] = 0$ the key is deleted from $D$. The whole insertion procedure is therefore performed in constant time.

@@

	The size of the queue is $w\log_2 |\Gamma|$, the size of the dictionary is $w (\log_2 |\Gamma| + \log_2 w)$ (as the dictionary cannot have more than $w$ keys at the same time, a dictionary key occupies $\log_2 |\Gamma|$ bits and a counter cannot go over $w$, thus being less than $\log_2 w$ bits long). Thus a requirement of $w (\log_2(w) + 2\log_2(|\Gamma|))$ bits for this DDF to work.

@@

	On the other hand, $e^\star \dup E$ is classified as \UNSEEN if and only if it was not seen in the last $M$ elements of the stream, and the filter answers \UNSEEN. Let $D$ be the event \enquote{\emph{There is at least one duplicate in the stream}} and $C$ be the event \enquote{\emph{There is a duplicate of $e^\star$ in the $M$ previous elements of the stream}}. Then $e^\star$ triggers a false negative with probability
	\begin{align*}
		\FN_n &= (1 - \Pr[C | D])(1-p) =  \left(1 - \frac{\Pr[C \cap D]}{\Pr[D]}\right)(1-p)\\
		& = \left(1 - \frac{\Pr[C]}{\Pr[D]}\right)(1-p) =  \left(1 - \frac{1 - \Pr[\bar C]}{1 - \Pr[\bar D]}\right)(1-p) \\
		\FN_n &=  \left(1 - \frac{1 - \left(1 - \frac 1{|\Gamma|}\right)^M}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n}\right)(1-p)
	\end{align*}
	Hence, the error probability of the filter is
	\begin{align*}
		E_n 
		& = \FN_n + p
		=  \left(1 - \frac{1 - \left(1 - \frac{1}{|\Gamma|}\right)^{M}}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n}\right)(1-p) + p
		= 1 - \frac{1-\left(1 - \frac{1}{|\Gamma|}\right)^{M}}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n}(1-p),
	\end{align*}
	which is minimized when $p = 0$. 
\qed\end{proof}

@@

	Each subfilter has undergone $c$ insertions, safe for the first subfilter which has only undergone $m \bmod c$, we immediately get Eq.~(\ref{eq:queue-pfp}).
\qed\end{proof}

@@

Finally, each event is weighted by the probability $p_c$ that $e^\star$ is in $\mathcal F_i$'s sub-sliding window:
\begin{align*}
	p_c 
	& = \Pr[\text{$e^\star$ is in }\mathcal F_i \text{ sub-sliding window | } e^\star \wdup E]\\
	& = \frac{\Pr[\text{$e^\star$ is in }\mathcal F_i \text{ sub-sliding window } \cap e^\star \wdup E]}{\Pr[e^\star \wdup E]}\\
	& = \frac{\Pr[\text{$e^\star$ is in }\mathcal F_i \text{ sub-sliding window}]}{\Pr[e^\star \wdup E]}\\
%		\end{align*}\begin{align*}p_c
	& = \frac{1 - \Pr[\text{$e^\star$ is not in }\mathcal F_i \text{ sub-sliding window }]}{1 - \Pr[e^\star \notwdup E]}\\
	& = \frac{1 - \left(1 - \frac 1{|\Gamma|}\right)^c}{1 - \left(1-\frac{1}{|\Gamma|}\right)^w}
\end{align*}
This concludes the proof.
\qed\end{proof}

@@

\begin{proof}
	If $cL \leq w$, then information-theoretically the subfilters only have information on elements in the sliding window.
	The probability of false positive for $\mathcal Q$ is $1 - (1 - \FP_{\mathcal F, c})^L$, which is strictly increasing with $\FP_{\mathcal F,c}$. Hence, the optimal solution is reached by to maximising the probability of false positive in each subfilter $\mathcal F_i$. By hypothesis the latter is bounded above by $p$ after $c$ insertions.

	On the other hand, if $cL > w$ then oldest filter holds information about elements that are not in the sliding window anymore. Hence, a strategy for the attacker trying to trigger a false positive on $e^\star$ could be to make it so these oldest elements are all equal to $e^\star$. Let $E$ be the optimal adversarial stream for triggering a false positive on the sliding window $w$ with the element $e^\star$, when $cL \leq w$. The adversary $\mathcal A$ can create a new stream $E' = e^\star | e^\star | \dotsc | E$ where $e^\star$ is prepended $cL - w$ times to $E$.

	After $w$ insertions, the last subfilter will answer $\DUPLICATE$ with probability at least $p$, hence giving a lower bound on $\mathcal A's$ probability of success. If, for some reason, the last subfilter answers $\DUPLICATE$ with probability less than $p$, then the same reasoning as for when $cL \leq w$ still applies, hence we get the correspondig lower bound (which is, in this case, an equality).
\qed\end{proof}

@@

\begin{proof}
	Let us first prove that a PPT adversary $\mathcal A$ can win the game with probability at least $p^L$.
	For this, let us consider the adversarial game against the subfilter $\mathcal F$: after $c$ insertions from an aversarial stream $E_c$, $\mathcal A$ choses a duplicate $e^\star$ which will be a false negative with proability $p$.
	Hence, if $\mathcal A$ crafts, for the filter $\mathcal Q$, the following adversarial stream $E' = E_c \mid E_c \mid \cdots \mid E_c$ consisting of $L$ concatenations of the stream $E_c$, then $e^\star$ is a false negative for $\mathcal Q$ if and only if it is a false negative for all subfilters $\mathcal F_i$, hence a probability of success for $\mathcal A$ of $p^L$.

	Now, Let us prove the case where $w\leq (L-1)c$. In this case, at any time, $\mathcal Q$ remembers all elements from inside the sliding window. As we have seen in the previous example, the probability of success of $\mathcal A$ is strictly increasing with the probability of each subfilter to answer $\UNSEEN$. The probability of a subfilter to answer $\UNSEEN$ is:

>>>>

\section{Compact Hash Short Filter}\label{app:CSHF}
Removing the dictionary from the SHF construction yields a more memory-efficient, but less time-efficient construction, which we dub \enquote{compact} short hash filter (CSHF). The CSHF performs in linear time in $w$, and is a simple queue, the only point is that instead of storing $e$, one stores $h(e)$, where $h$ is a hash function of codomain size $\lfloor \frac Mw \rfloor$.% and is described in Algorithm~\ref{alg:CSHF}.

%\begin{algorithm}[!h]
%	\caption{CSHF, Setup, Lookup and Insert}\label{alg:CSHF}
%	
%	\begin{algorithmic}[1]
%		\Function{Setup}{$\mathcal M, w$}
%		\Comment $M$ is the available memory, $w$ the size of the sliding window
%		\State $h \gets$ hash function of codomain size $\lfloor \frac Mw \rfloor$
%		\State $Q \gets \emptyset$ \Comment $Q$ is a queue of elements of size $h$
%		\State\Return filter
%		\EndFunction 
%		\item[]
%	\end{algorithmic}
%	
%	\begin{algorithmic}[1]
%		\Function{Lookup}{$e$}
%		\State \Return $h(e) \in Q$ 
%		\EndFunction
%		\item[]
%	\end{algorithmic}
%	
%	\begin{algorithmic}[1]
%		\Function{Insert}{$e$}
%		\State $Q.\mathsf{Push\_Front}(h(e))$
%		
%		\If{$Q.\mathsf{length}() > w$}
%			\State $h' \gets Q.\mathsf{Pop\_back}()$
%		\EndIf
%		\EndFunction
%	\end{algorithmic}
%\end{algorithm}

\paragraph{Error Rate.}
Let $w> 0$ be a window size and $M > 0$ the available memory.

%We write $\FN^w_\text{SHF}$ the probability of false negative of an SHF with these parameters. We similarly define $\FP^w_\text{SHF}$.%, $\FN^w_\text{CSHF}$, $\FP^w_\text{CSHF}$.

\begin{restatable}{thm}{fprSHF}\label{thm:fprCHF}	
	$\FN_\text{CSHF}^w  = 0$  and $\FP_\text{CSHF}^w = 1 - \left(1 - 2^{-M/w}\right)^w$
\end{restatable}

\begin{proof}
	This is an immediate adaptation of the proof from Theorem~\ref{thm:SHF}. A CSHF has fingerprints of size $h' = \frac Mw$.
\end{proof}

\paragraph{Remark:} A CSHF of size $M$ has the same error rate than an SHF of size $2M + w\log_2 w$.

\paragraph{Saturation.}
Simlarly to SHF, CSHF saturates quickly over a certain $w_\text{max}$. We compare SHF and CSHF saturations in Figure~\ref{fig:app_shorts}.

\begin{figure}[t]
		\centering
		\input{graphs/bench_shorts_both.tex}
		\caption{Error rates of SHFs and CSHFs for $M = 10^5$ bits, for varying window sizes $w$.}\label{fig:app_shorts}
\end{figure}

Experimentally, we get $w_\text{max}^\text{CSHF} = 0.0627 M + 443$ ($r^2 = 0.9981$).

@@

	When an element $e$ is inserted in the DDF, $e$ is stored and $D[e]$ is incremented (if the key $e$ did not exist in $D$, it is created first, and
	$D[e]$ is set to $1$). In order to keep the number of stored elements to $w$, 
	we discard the oldest element $e_\text{last}$ in $Q$. As we do so, we also decrement $D[e_\text{last}]$, and if $D[e_{\text{last}}] = 0$ the key is deleted from $D$. The whole insertion procedure is therefore performed in constant time.

@@

	The queue  size is $w\log_2 |\Gamma|$, the dictionary size is $w (\log_2 |\Gamma| + \log_2 w)$ (as the dictionary cannot have more than $w$ keys at the same time, a dictionary key occupies $\log_2 |\Gamma|$ bits and a counter cannot go over $w$, thus being less than $\log_2 w$ bits long). Thus a requirement of $w (\log_2(w) + 2\log_2(|\Gamma|))$ bits for this DDF to work.

@@

	On the other hand, $e^\star \dup E$ is classified as \UNSEEN if and only if it was not seen in the last $M$ elements of the stream, and the filter answers \UNSEEN. Let $D$ be the event \enquote{\emph{There is at least one duplicate in the stream}} and $C$ be the event \enquote{\emph{There is a duplicate of $e^\star$ in the $M$ previous elements of the stream}}. Then $e^\star$ triggers a false negative with probability
	\begin{align*}
		\FN_n &= (1 - \Pr[C | D])(1-p) =  \left(1 - \frac{\Pr[C \cap D]}{\Pr[D]}\right)(1-p)
		 = \left(1 - \frac{\Pr[C]}{\Pr[D]}\right)(1-p) \\
		\FN_n &=  \left(1 - \frac{1 - \Pr[\bar C]}{1 - \Pr[\bar D]}\right)(1-p)
			=  \left(1 - \frac{1 - \left(1 - \frac 1{|\Gamma|}\right)^M}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n}\right)(1-p)
	\end{align*}
	Hence, the error probability of the filter is
	\begin{align*}
		EP_n 
		& = \FN_n + p
		=  \left(1 - \frac{1 - \left(1 - \frac{1}{|\Gamma|}\right)^{M}}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n}\right)(1-p) + p
		= 1 - \frac{1-\left(1 - \frac{1}{|\Gamma|}\right)^{M}}{1 - \left(1 - \frac 1{|\Gamma|}\right)^n}(1-p),
	\end{align*}
	which is minimized when $p = 0$. 
\qed\end{proof}

@@

	Each subfilter has undergone $c$ insertions, except for the first subfilter which has only undergone $m \bmod c$, we immediately get Eq.~(\ref{eq:queue-pfp}).
\qed\end{proof}

@@

Finally, each event is weighted by the probability $p_c$ that $e^\star$ is in $\mathcal F_i$'s sub-sliding window:
\begin{align*}
	p_c 
	& = \Pr[\text{$e^\star$ is in }\mathcal F_i \text{ sub-sliding window | } e^\star \wdup E]
	 = \frac{\Pr[\text{$e^\star$ is in }\mathcal F_i \text{ sub-sliding window } \cap e^\star \wdup E]}{\Pr[e^\star \wdup E]}\\
	& = \frac{\Pr[\text{$e^\star$ is in }\mathcal F_i \text{ sub-sliding window}]}{\Pr[e^\star \wdup E]}
%		\end{align*}\begin{align*}p_c
	 = \frac{1 - \Pr[\text{$e^\star$ is not in }\mathcal F_i \text{ sub-sliding window }]}{1 - \Pr[e^\star \notwdup E]}\\
	p_c& = \frac{1 - \left(1 - \frac 1{|\Gamma|}\right)^c}{1 - \left(1-\frac{1}{|\Gamma|}\right)^w}
\end{align*}
This concludes the proof.
\qed\end{proof}

@@

\begin{proof}
	If $cL \leq w$, then information-theoretically the subfilters only have information on elements in the sliding window.
	The false positive probability for $\mathcal Q$ is $1 - (1 - \FP_{\mathcal F, c})^L$, which is strictly increasing with $\FP_{\mathcal F,c}$. Hence, the optimal solution is reached by to maximising the false positive probability in each subfilter $\mathcal F_i$. By hypothesis the latter is bounded above by $p$ after $c$ insertions.

	On the other hand, if $cL > w$ then the oldest filter holds information about elements that are not in the sliding window anymore. Hence, a strategy for the attacker trying to trigger a false positive on $e^\star$ could be to make it so these oldest elements are all equal to $e^\star$. Let $E$ be the optimal adversarial stream for triggering a false positive on the sliding window $w$ with the element $e^\star$, when $cL \leq w$. The adversary $\mathcal A$ can create a new stream $E' = e^\star | e^\star | \dotsc | E$ where $e^\star$ is prepended $cL - w$ times to $E$.

	After $w$ insertions, the last subfilter will answer $\DUPLICATE$ with probability at least $p$, hence giving a lower bound on $\mathcal A's$ success probability. If, for some reason, the last subfilter answers $\DUPLICATE$ with probability less than $p$, then the same reasoning as for when $cL \leq w$ still applies, hence we get the correspondig lower bound (which is, in this case, an equality).
\qed\end{proof}

@@

\begin{proof}
	Let us first prove that a PPT adversary $\mathcal A$ can win the game with probability at least $p^L$.
	For this, let us consider the adversarial game against the subfilter $\mathcal F$: after $c$ insertions from an aversarial stream $E_c$, $\mathcal A$ choses a duplicate $e^\star$ which will be a false negative with proability $p$.
	Hence, if $\mathcal A$ crafts, for the filter $\mathcal Q$, the following adversarial stream $E' = E_c \mid E_c \mid \cdots \mid E_c$ consisting of $L$ concatenations of the stream $E_c$, then $e^\star$ is a false negative for $\mathcal Q$ if and only if it is a false negative for all subfilters $\mathcal F_i$, hence a success probability for $\mathcal A$ of $p^L$.

	Now, Let us prove the case where $w\leq (L-1)c$. In this case, at any time, $\mathcal Q$ remembers all elements from inside the sliding window. As we have seen in the previous example, the success probability of $\mathcal A$ is strictly increasing with the probability of each subfilter to answer $\UNSEEN$. The probability of a subfilter to answer $\UNSEEN$ is:
